---
title       : "Causal impact of masks, policies, behavior on early COVID-19 pandemic in the U.S."
subtitle    : "Counterfactuals"
author      :  Victor Chernozhukov, Paul Schrimpf, Hiro Kasahara
job         :
date        : "`r format(Sys.time(), '%d %B, %Y')`"
bibliography: "covid.bib"
link-citations: true
always_allow_html: true
output      :
    html_document :
        toc : true
        toc_depth : 3
        toc_float : true
        number_sections : true
        #theme : journal
        #css : 628notes.css
        code_folding: hide
        #lib_dir : deps
        self_cononontained : true
        fig_width: 8
        fig_height: 6
---

 <!-- To create html files from this, in R enter  -->
 <!-- library(rmarkdown) -->
 <!-- rmarkdown::render("regressionsWithPlots.Rmd") -->
 <!-- To create tex files from this, in R enter ` -->
 <!-- library(rmarkdown) -->
 <!-- rmarkdown::render("regressionsWithPlots.Rmd", output_format=latex_fragment()) -->


 <!-- to create the gh-pages hosted website in R enter -->
 <!-- rmarkdown::render_site() -->

 <!-- Then in a shell  -->
 <!-- $ ghp-import -p -n -m "$(date)" _site -->


```{r, include=FALSE}
options(knitr.table.format = function() {
  if (knitr::is_latex_output()) 'latex' else 'pandoc'
})
if (knitr::is_latex_output()) {
  knitr::opts_chunk$set(warning=FALSE, message=FALSE, echo=FALSE)
}
```

# Data Preparation

```{r setup, cache=FALSE, warning=FALSE, include=FALSE}

library(lfe)
library(stargazer)
library(knitr)
library(plm)
library(latex2exp)
library(ggplot2)
library(ggthemes)
library(estimatr)
library(gridExtra)
library(grid)
library(mvtnorm)
library(kableExtra)
library(data.table)
library(reshape2)

# stupid tidyverse stealing base package names ...
if("dplyr" %in% (.packages())){
  detach("package:dplyr", unload=TRUE)
}

sgtype <- opts_knit$get("rmarkdown.pandoc.to")
sgstyle <- 'default'
colors <-  scale_color_solarized
colors_fill <- scale_fill_solarized
rootdir <- system("git rev-parse --show-toplevel", intern=TRUE)[1]
source(paste(rootdir,"cases_and_policies/R/dataprep.R",sep="/"))
source(paste(rootdir,"cases_and_policies/rmd/varlabels.R", sep="/"))
source(paste(rootdir,"cases_and_policies/R/utils.R", sep="/"))
source(paste(rootdir,"cases_and_policies/rmd/regprep.R", sep="/"))
source(paste(rootdir,"cases_and_policies/rmd/generatetables.R", sep="/"))
source(paste(rootdir,"cases_and_policies/rmd/bootstrap_felm.R", sep="/"))

datelims <- c(as.Date("2020-03-07"), max(df$date))
```

# Counterfactuals


```{r cffuncs, cache=FALSE}
source(paste(rootdir,"cases_and_policies/rmd/counterfactuals.R", sep="/"))
```

```{r statesim, cache=FALSE}
infovars <- list(c("dlogdc", "logdc")) #,
#c("dlogdc", "logdc"))
#infovars2 <- list(c( "logdc"))
tvars <- "dlogtests"
xlist <-list("")
interactions <-list("month", statevars)
ilist <- list(interactions) #, interactions)
iv <- list("0") #,"(testratedc ~ lag(testratedc,14))")

L <- 14
sdf <- subset(df, as.vector(df$date)>=datelims[1]) #+L)
regs <- mainregressions(sdf,
                         "dlogdc", pols, bvars, infovars, tvars, xlist, ilist, iv, L=L)
m <- regs$piy[[1]]$reg
cnames <- c(pols, infovars[[1]])
di <- dieff_table(lapply(regs$pib[[1]][1:4], function(x) x$reg$coef[,1]),
                   regs$pbiy[[1]]$reg$coef[,1],
                   regs$piy[[1]]$reg$coef[,1],
                   policies=cnames)

pib <- lapply(regs$pib[[1]][1:4], function(x) x$reg)
pbiy <- regs$pbiy[[1]]$reg
piy <- regs$piy[[1]]$reg

S <- 999
bs <- bootstrap_felm(sdf, c(pib,list(pbiy,piy)), S=S)
bdi <- lapply(1:S, function(i)
  dieff_table(lapply(1:4, function(j) bs[[j]][,i]),
              bs[[5]][,i], bs[[6]][,i],
              policies=cnames)
  )

subcoefs <- function(di, coef) {
  for (r in 1:(nrow(di)-1)) {
    i <- grep(rownames(di)[r] , rownames(coef))
    coef[i,1] <- di[r,"Average"]
  }
  return(coef)
}

bcoefs <- t(do.call(cbind, lapply(1:S, function(i) {
  coef <- matrix(bs[[6]][,i],ncol=1)
  rownames(coef) <- rownames(bs[[6]])
  return(subcoefs(bdi[[i]],coef ))
})))

V <- cov(bcoefs)
m$coefficients <- subcoefs(di, m$coefficients)
m$vcv <- V
m$clustervcv <- V
m$robustvcv <- V

simmodel <- createforwardsim(m, sdf)

# remove policies
cf <- sdf
cf[,pols] <- 0
nopmodel <- createforwardsim(m, cf)

rerunsim=FALSE
if (!file.exists("cfnop.Rda") || rerunsim) {
  cfnoplist <- lapply(unique(sdf$state), function(s)
    nopstatesim(s,simobs=simmodel,simcf=nopmodel,data=sdf))
  allmeans <- rbindlist(lapply(cfnoplist, getmeans))
  save(cfnoplist,allmeans, file="cfnop.Rda")
}
load("cfnop.Rda")
cfnop <- do.call(rbind, lapply(cfnoplist, function(x) x$edf))
```

### Removing Policies

We now examine the impact changing from the observed policies to
none. For illustrative purpose, we being by focusing on
Washington. Figure \ref{fig:WA-nop-dlogdc} shows the observed,
estimated average, and counterfactual without policies average of
$\Delta \log \Delta C$ in Washington.

We hold residuals fixed throughout.
<!-- To compute the estimated and counterfactual paths we use the estimates -->
<!-- from column 2 of table \ref{tab:policy-cases}. We set initial $\Delta -->
<!-- \log \Delta C$ and $\log \Delta C$ to their values first observed in -->
<!-- the state we are simulating. We hold all other regressors at their -->
<!-- observed values. Error terms are drawn with replacement from the -->
<!-- residuals. We do this many times and report the average over draws of -->
<!-- the residuals. To calculate confidence intervals, we repeat the above -->
<!-- with coefficients drawn randomly from their asymptotic -->
<!-- distribution. The shaded region is a pointwise 90\% confidence -->
<!-- interval. -->

<!-- To be more precise, we can rewrite our regression model as: -->

<!-- $$ -->
<!--   Y_{it} = P_{it}'\alpha + \sum_{\ell=1}^L \rho_\ell Y_{it-\ell} + X_{it}'\beta + \epsilon_{it} -->
<!-- $$ -->

<!-- where $Y_{it} = \log \Delta C_{it}$, $P_{it}$ are policies, and -->
<!-- $X_{it}$ are controls. Each counterfactual involves changing $P_{it}$ -->
<!-- to some other value, $P_{it}^\ast$. We then generate counterfactual paths -->
<!-- for cases by in state $i$ conditional on -->
<!-- $Y_{i1}, ..., Y_{iL}, \{X_{it}, P_{it}^\ast \}_{t=1}^T, \theta=(\alpha, \beta, \rho)$ by setting: -->

<!-- $$ -->
<!--   Y_{it}^{\ast,r} = {P_{it}^\ast}'\alpha + \sum_{\ell=1}^L \rho_\ell Y_{it-\ell} + X_{it}'\beta + \hat{\epsilon}_{it}^{\ast,r} -->
<!-- $$ -->

<!-- where $\hat{\epsilon}_{it}^{\ast,r}$ are resampled with replacement -->
<!-- from $\{\hat{\epsilon}_{it}\}_{i=1,t=1}^{N,T}$. We do resample the -->
<!-- residuals $R$ times, repeating the above each time. We then compute -->

<!-- $$ -->
<!-- \hat{E}[Y_{it}^\ast | Y_{i1}, ..., Y_{iL}, \{X_{it}, P_{it}^\ast\}_{t=1}^T, \theta] = -->
<!-- \frac{1}{R} \sum_{r=1}^R Y_{it}^{\ast,r} -->
<!-- $$ -->

<!-- Finally, we repeatedly draw $\theta$ from its asymptotic distribution -->
<!-- and repeat all of the above to compute confidence intervals. -->

<!-- For case growth, the same procedure is used, but with $\Delta Y_{it} = \Delta \log \Delta C_{it}$ in place of $Y_{it}$. The top panel of the following figure shows -->

<!-- $$ -->
<!-- E[\Delta Y_{it} | Y_{i1}, ..., Y_{iL}, \{X_{it}, P_{it}\}_{t=1}^T] -->
<!-- $$ -->

<!-- and -->

<!-- $$ -->
<!-- E[\Delta Y_{it}^\ast | Y_{i1}, ..., Y_{iL}, \{X_{it}, P_{it}^\ast\}_{t=1}^T] -->
<!-- $$ -->

<!-- along with the observed $Y_{it}$, and the bottom panel shows the difference. -->


<!-- In this subsection, instead of resampling residuals, we condition on them. That is we look at -->

<!-- $$ -->
<!--   Y_{it}^{\ast} = {P_{it}^\ast}'\alpha + \sum_{\ell=1}^L \rho_\ell Y_{it-\ell} + X_{it}'\beta + \hat{\epsilon}_{it} -->
<!-- $$ -->

<!-- instead of -->

<!-- $$ -->
<!--   \frac{1}{R} \sum_{r=1}^R Y_{it}^{\ast,r}. -->
<!-- $$ -->


```{r WA-nop-dlogdc, cache=TRUE,fig.cap="Case growth without policies in Washington \\label{fig:WA-nop-dlogdc}"}
st <- "Washington"
f <- cfplots(st, cfnop, data=df)
grid.arrange(f$figd, f$figdp, nrow=2, top =
                                        textGrob(st,gp=gpar(fontsize=20,font=2)))

```

To look at cases we instead compute,

$$
\hat{E}[\exp(Y_{it}^\ast) | Y_{i1}, ..., Y_{iL}, \{X_{it}, P_{it}^\ast\}_{t=1}^T, \theta] =
\frac{1}{R} \sum_{r=1}^R \exp(Y_{it}^{\ast,r})
$$


```{r WA-nop-cases, cache=TRUE, fig.cap="Cases without policies in Washington \\label{fig:WA-nop-cases}"}
grid.arrange(f$figl, f$figp, nrow=2) #,
#top = textGrob(st,gp=gpar(fontsize=20,font=2)))
```

## National

These figures show the average effect on $\Delta \log \Delta C$ across
states, and the total national effect on cases in past week.

That is for the first figure, we take:

$$
\hat{E}[\Delta Y_{it}^\ast | Y_{i1}, ..., Y_{iL}, \{X_{it}, P_{it}^\ast\}_{t=1}^T, \theta] =
\frac{1}{N} \sum_{i=1}^{N} \frac{1}{R} \sum_{r=1}^R \Delta Y_{it}^{\ast,r}
$$

and plot the difference of the above between two different values of $P_{it}^\ast$

The second figure shows the total change in cases, that is the change in:

$$
\hat{E}[\exp(Y_{it}^\ast) | Y_{i1}, ..., Y_{iL}, \{X_{it}, P_{it}^\ast\}_{t=1}^T, \theta] =
\sum_{i=1}^{N} \frac{1}{R} \sum_{r=1}^R \exp(Y_{it}^{\ast,r}).
$$

The final figure shows
$$
\frac{\sum_{i=1}^{N} \frac{1}{R} \sum_{r=1}^R \exp(Y_{it}^{\ast,r}) - \sum_{i=1}^{N} \frac{1}{R} \sum_{r=1}^R \exp(Y_{it}^r)} {\sum_{i=1}^{N} \frac{1}{R} \sum_{r=1}^R \exp(Y_{it}^{r})}
$$
that is, the (average over residual draws) change in national cases relative to (average over residual draws) national cases.


```{r nop-all, cache=TRUE}
abbrv <- unique(df[,c("state","ST")])
cfnop <- merge(cfnop, abbrv, by="state")

figs <- nationalplots(allmeans,cfnop, "removing policies")

figdev(paste(rootdir,"tex/tables_and_figures/us-nop-dgrowth_v1.pdf",sep="/"), width=5,height=3.75)
figs$g
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-nop-dcases_v1.pdf",sep="/"), width=5,height=3.75)
figs$c
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-nop-rel_v1.pdf",sep="/"), width=5,height=3.75)
figs$r
dev.off()

for (f in figs) print(f)
```

## Non-Essential Business Closures

```{r business_sim, cache=FALSE}
# remove non essential business closure
cf <- sdf
cf[,"pnonessential"] <- 0
nobmodel <- createforwardsim(m, cf)

if (!file.exists("cfnb.Rda") || rerunsim) {
  cfnblist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=nobmodel, data=sdf, nresid=1))
  nball <- rbindlist(lapply(cfnblist, getmeans))
  save(cfnblist,nball, file="cfnb.Rda")
}
load("cfnb.Rda")
cfnb <- do.call(rbind, lapply(cfnblist, function(x) x$edf))
```

```{r, cache=TRUE}
cfnb <- merge(cfnb, abbrv, by="state")
for (st in c("Washington","Illinois","Massachusetts")) {
  f <- cfplots(st, cfnb, data=df)
  f$figp + ggtitle(sprintf("Effect of never shutting down non-essential businesses in %s",st))
  #grid.arrange(f$figd + ggtitle(), f$figdp + ggtitle(), nrow=2, top =
  #textGrob(st,gp=gpar(fontsize=20,font=2)))

  figdev(sprintf("%s/tex/tables_and_figures/%s-nb-dgrowth_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figdp)
  dev.off()


  figdev(sprintf("%s/tex/tables_and_figures/%s-nb-dcases_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figp)
  dev.off()
}


figs <- nationalplots(nball, cfnb,
                      "never shutting down non-essential\nbusinesses")

figdev(paste(rootdir,"tex/tables_and_figures/us-nb-dgrowth_v1.pdf",sep="/"), width=5,height=3.75)
figs$g
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-nb-dcases_v1.pdf",sep="/"), width=5,height=3.75)
figs$c
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-nb-rel_v1.pdf",sep="/"), width=5,height=3.75)
figs$r
dev.off()

for (f in figs) print(f)
```

## Masks

We now look at what happens when all states impose a mask mandate on
April 1st.

```{r mask_sim, cache=FALSE}
cf <- df
cf[,"Mandate.face.mask.use.by.employees.in.public.facing.businesses"]  <- as.Date("2020-04-01")
cf[,"pmaskbus"] <-
  smoothedpolicy(cf,"Mandate.face.mask.use.by.employees.in.public.facing.businesses",type="ma",bw=6)
cf <- subset(cf, cf$date>=datelims[1]) #+L)
maskmodel <- createforwardsim(m, cf)

if (!file.exists("cfmask.Rda") || rerunsim) {
  cfmasklist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=maskmodel, data=sdf, nresid=1))
  maskall <- rbindlist(lapply(cfmasklist, getmeans))
  save(cfmasklist, maskall, file="cfmask.Rda")
}
load("cfmask.Rda")
cfmask <- do.call(rbind, lapply(cfmasklist, function(x) x$edf))
```

```{r, cache=TRUE}
cfmask <- merge(cfmask, abbrv, by="state")
for (st in c("Washington","Illinois","Massachusetts")) {
  f <- cfplots(st, cfmask, data=df)
  print(f$figp +  ggtitle(sprintf("Effect of mandating masks on April 1st in %s",st)))

  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-dgrowth_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figdp)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-dcases_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figp)
  dev.off()
}

figs <- nationalplots(maskall, cfmask, "mandating masks on April 1st\n")

figdev(paste(rootdir,"tex/tables_and_figures/us-mask-dgrowth_v1.pdf",sep="/"), width=5,height=3.75)
figs$g
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-mask-dcases_v1.pdf",sep="/"), width=5,height=3.75)
figs$c
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-mask-rel_v1.pdf",sep="/"), width=5,height=3.75)
figs$r
dev.off()

for (f in figs) print(f)
```



## Stay at home

We now look at what happens if stay at home orders were never issued.

```{r home, cache=FALSE}
cf <- df
cf[,"pshelter"] <- 0
cf <- subset(cf, cf$date>=datelims[1]) #+L)
noshelter <- createforwardsim(m, cf)

if (!file.exists("cfshelter.Rda") || rerunsim) {
  cfshelterlist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=noshelter, data=sdf, nresid=1))
  shelterall <- rbindlist(lapply(cfshelterlist, getmeans))
  save(cfshelterlist, shelterall, file="cfshelter.Rda")
}
load("cfshelter.Rda")
cfshelter <- do.call(rbind, lapply(cfshelterlist, function(x) x$edf))
```

```{r, cache=TRUE}
cfshelter <- merge(cfshelter, abbrv, by="state")
for (st in c("Washington","Illinois","Massachusetts")) {
  f <- cfplots(st, cfshelter, data=df)
  print(f$figp +  ggtitle(sprintf("Effect of no stay at home order in %s",st)))

  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dgrowth_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figdp)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dcases_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figp)
  dev.off()
}

figs <- nationalplots(shelterall, cfshelter, "no stay at home orders")

figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-dgrowth_v1.pdf",sep="/"), width=5,height=3.75)
figs$g
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-dcases_v1.pdf",sep="/"), width=5,height=3.75)
figs$c
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-rel_v1.pdf",sep="/"), width=5,height=3.75)
figs$r
dev.off()

for (f in figs) print(f)
```

## Shutdown policies

We now look at what happens if the four shutdown policies in pindex were never implemented.

```{r shutdown, cache=FALSE}
cf <- df
cf[,"pshelter"] <- 0
cf[,"pmovie"] <- 0
cf[,"prestaurant"] <- 0
cf[,"pnonessential"] <- 0
cf <- subset(cf, cf$date>=datelims[1]) #+L)
noindex <- createforwardsim(m, cf)

if (!file.exists("cfindex.Rda") || rerunsim) {
  cfindexlist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=noindex, data=sdf, nresid=1))
  indexall <- rbindlist(lapply(cfindexlist, getmeans))
  save(cfindexlist, indexall, file="cfindex.Rda")
}
load("cfindex.Rda")
cfindex <- do.call(rbind, lapply(cfindexlist, function(x) x$edf))
```

```{r, cache=TRUE}
cfindex <- merge(cfindex, abbrv, by="state")
for (st in c("Washington","Illinois","Massachusetts")) {
  f <- cfplots(st, cfindex, data=df, yvar="deaths")
  print(f$figp +  ggtitle(sprintf("Effect of no shutdown policies in %s",st)))

  figdev(sprintf("%s/tex/tables_and_figures/%s-index-dgrowth.pdf",rootdir,st), width=5,height=3.75)
  print(f$figdp)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-index-dcases.pdf",rootdir,st), width=5,height=3.75)
  print(f$figp)
  dev.off()
}

figs <- nationalplots(indexall, cfindex, "no shutdown policies", yvar="deaths")

figdev(paste(rootdir,"tex/tables_and_figures/us-index-dgrowth.pdf",sep="/"), width=5,height=3.75)
figs$g
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-index-dcases.pdf",sep="/"), width=5,height=3.75)
figs$c
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-index-rel.pdf",sep="/"), width=5,height=3.75)
figs$r
dev.off()

for (f in figs) print(f)
```


# Deaths

```{r statesim-deaths, cache=FALSE}
infovars <- list(c("dlogdd", "logdd")) #,
#c("dlogdc", "logdc"))
#infovars2 <- list(c( "logdc"))
tvars <- NULL
xlist <-list("")
interactions <-list("month", statevars)
ilist <- list(interactions) #, interactions)
iv <- list("0") #,"(testratedc ~ lag(testratedc,14))")

L <- 21
sdf <- subset(df, df$date>=datelims[1]) #+L)
regs <- mainregressions(sdf,
                        "dlogdd", pols, bvars, infovars, tvars, xlist, ilist, iv, L=L)

m <- regs$piy[[1]]$reg

cnames <- c(pols, infovars[[1]])
di <- dieff_table(lapply(regs$pib[[1]][1:4], function(x) x$reg$coef[,1]),
                  regs$pbiy[[1]]$reg$coef[,1],
                  regs$piy[[1]]$reg$coef[,1],
                  policies=cnames)

pib <- lapply(regs$pib[[1]][1:4], function(x) x$reg)
pbiy <- regs$pbiy[[1]]$reg
piy <- regs$piy[[1]]$reg

S <- 999
bs <- bootstrap_felm(sdf, c(pib,list(pbiy,piy)), S=S)
bdi <- lapply(1:S, function(i)
  dieff_table(lapply(1:4, function(j) bs[[j]][,i]),
              bs[[5]][,i], bs[[6]][,i],
              policies=cnames)
  )

bcoefs <- t(do.call(cbind, lapply(1:S, function(i) {
  coef <- matrix(bs[[6]][,i],ncol=1)
  rownames(coef) <- rownames(bs[[6]])
  return(subcoefs(bdi[[i]],coef ))
})))
V <- cov(bcoefs)
m$coefficients <- subcoefs(di, m$coefficients)
m$vcv <- V
m$clustervcv <- V
m$robustvcv <- V

simmodel <- createforwardsim(m, sdf, dvar="dlogdd", var="logdd")

# remove policies
cf <- sdf
cf[,pols] <- 0
nopmodel <- createforwardsim(m, cf, dvar="dlogdd", var="logdd")

rerunsim=FALSE
if (!file.exists("cfnop_deaths.Rda") || rerunsim) {
  cfnoplist <- lapply(unique(df$state), function(s)
    nopstatesim(s,simobs=simmodel,simcf=nopmodel, data=sdf, var="logdd", dvar="dlogdd"))
  allmeans <- rbindlist(lapply(cfnoplist, getmeans))
  save(cfnoplist,allmeans, file="cfnop_deaths.Rda")
}
load("cfnop_deaths.Rda")
cfnop <- do.call(rbind, lapply(cfnoplist, function(x) x$edf))
```

## Removing Policies


```{r WA-nop-dlogdd, cache=TRUE,fig.cap="Case growth without policies in Washington \\label{fig:WA-nop-dlogdc}"}
st <- "Washington"
f <- cfplots(st, cfnop, yvar="deaths", data=df)
grid.arrange(f$figd, f$figdp, nrow=2, top =
                                        textGrob(st,gp=gpar(fontsize=20,font=2)))

```

```{r WA-nop-deaths, cache=TRUE, fig.cap="Cases without policies in Washington \\label{fig:WA-nop-cases}"}
grid.arrange(f$figl, f$figp, nrow=2) #,
#top = textGrob(st,gp=gpar(fontsize=20,font=2)))
```

### National


```{r nop-all-deaths, cache=TRUE}
abbrv <- unique(df[,c("state","ST")])
cfnop <- merge(cfnop, abbrv, by="state")

figs <- nationalplots(allmeans,cfnop, "removing policies", yvar="deaths")

figdev(paste(rootdir,"tex/tables_and_figures/us-nop-dgrowth_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$g
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-nop-dcases_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$c
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-nop-rel_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$r
dev.off()

for (f in figs) print(f)
```

## Non-essential Business Closures

```{r business_sim-deaths, cache=FALSE}
# remove non essential business closure
cf <- sdf
cf[,"pnonessential"] <- 0
nobmodel <- createforwardsim(m, cf, dvar="dlogdd", var="logdd")

if (!file.exists("cfnb_deaths.Rda") || rerunsim) {
  cfnblist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=nobmodel, data=sdf, var="logdd", dvar="dlogdd"))
  nball <- rbindlist(lapply(cfnblist, getmeans))
  save(cfnblist,nball, file="cfnb_deaths.Rda")
}
load("cfnb_deaths.Rda")
cfnb <- do.call(rbind, lapply(cfnblist, function(x) x$edf))
```

```{r, cache=TRUE}
cfnb <- merge(cfnb, abbrv, by="state")
for (st in c("Washington","Illinois","Massachusetts")) {
  f <- cfplots(st, cfnb, yvar="deaths", data=df)
  f$figp + ggtitle(sprintf("Effect of never shutting down non-essential businesses in %s",st))
  #grid.arrange(f$figd + ggtitle(), f$figdp + ggtitle(), nrow=2, top =
  #textGrob(st,gp=gpar(fontsize=20,font=2)))

  figdev(sprintf("%s/tex/tables_and_figures/%s-nb-dgrowth_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figdp)
  dev.off()


  figdev(sprintf("%s/tex/tables_and_figures/%s-nb-dcases_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figp)
  dev.off()
}


figs <- nationalplots(nball, cfnb,
                      "never shutting down non-essential\nbusinesses", yvar="deaths")

figdev(paste(rootdir,"tex/tables_and_figures/us-nb-dgrowth_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$g
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-nb-dcases_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$c
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-nb-rel_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$r
dev.off()

for (f in figs) print(f)
```

## Masks

We now look at what happens when all states impose a mask mandate on
April 1st.

```{r mask_sim-deaths, cache=FALSE}
cf <- df
cf[,"Mandate.face.mask.use.by.employees.in.public.facing.businesses"]  <- as.Date("2020-04-01")
cf[,"pmaskbus"] <-
  smoothedpolicy(cf,"Mandate.face.mask.use.by.employees.in.public.facing.businesses",type="ma",bw=6)
cf <- subset(cf, cf$date>=datelims[1]) #+L)
maskmodel <- createforwardsim(m, cf, dvar="dlogdd", var="logdd")

if (!file.exists("cfmask_deaths.Rda") || rerunsim) {
  cfmasklist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=maskmodel, data=sdf, var="logdd", dvar="dlogdd"))
  maskall <- rbindlist(lapply(cfmasklist, getmeans))
  save(cfmasklist, maskall, file="cfmask_deaths.Rda")
}
load("cfmask_deaths.Rda")
cfmask <- do.call(rbind, lapply(cfmasklist, function(x) x$edf))
```

```{r, cache=TRUE}
cfmask <- merge(cfmask, abbrv, by="state")
for (st in c("Washington","Illinois","Massachusetts")) {
  f <- cfplots(st, cfmask, yvar="deaths", data=df)
  print(f$figp +  ggtitle(sprintf("Effect of mandating masks on April 1st in %s",st)))

  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-dgrowth_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figdp)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-dcases_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figp)
  dev.off()
}

figs <- nationalplots(maskall, cfmask, "mandating masks on April 1st\n", yvar="deaths")

figdev(paste(rootdir,"tex/tables_and_figures/us-mask-dgrowth_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$g
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-mask-dcases_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$c
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-mask-rel_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$r
dev.off()

for (f in figs) print(f)
```

## Stay at home

We now look at what happens if stay at home orders were never issued.

```{r home-deaths, cache=FALSE}
cf <- df
cf[,"pshelter"] <- 0
cf <- subset(cf, cf$date>=datelims[1]) #+L)
noshelter <- createforwardsim(m, cf, dvar="dlogdd", var="logdd")

if (!file.exists("cfshelter_deaths.Rda") || rerunsim) {
  cfshelterlist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=noshelter, data=sdf))
  shelterall <- rbindlist(lapply(cfshelterlist, getmeans))
  save(cfshelterlist, shelterall, file="cfshelter_deaths.Rda")
}
load("cfshelter_deaths.Rda")
cfshelter <- do.call(rbind, lapply(cfshelterlist, function(x) x$edf))
```

```{r, cache=TRUE}
cfshelter <- merge(cfshelter, abbrv, by="state")
for (st in c("Washington","Illinois","Massachusetts")) {
  f <- cfplots(st, cfshelter, data=df, yvar="deaths")
  print(f$figp +  ggtitle(sprintf("Effect of no stay at home order in %s",st)))

  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dgrowth_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figdp)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dcases_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figp)
  dev.off()
}

figs <- nationalplots(shelterall, cfshelter, "no stay at home orders", yvar="deaths")

figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-dgrowth_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$g
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-dcases_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$c
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-rel_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$r
dev.off()

for (f in figs) print(f)
```

## Shutdown policies

We now look at what happens if the four shutdown policies in pindex were never implemented.

```{r shutdown-deaths, cache=FALSE}
cf <- df
cf[,"pshelter"] <- 0
cf[,"pmovie"] <- 0
cf[,"prestaurant"] <- 0
cf[,"pnonessential"] <- 0
cf <- subset(cf, cf$date>=datelims[1]) #+L)
noindex <- createforwardsim(m, cf, dvar="dlogdd", var="logdd")

if (!file.exists("cfindex_deaths.Rda") || rerunsim) {
  cfindexlist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=noindex, data=sdf))
  indexall <- rbindlist(lapply(cfindexlist, getmeans))
  save(cfindexlist, indexall, file="cfindex_deaths.Rda")
}
load("cfindex_deaths.Rda")
cfindex <- do.call(rbind, lapply(cfindexlist, function(x) x$edf))
```

```{r, cache=TRUE}
cfindex <- merge(cfindex, abbrv, by="state")
for (st in c("Washington","Illinois","Massachusetts")) {
  f <- cfplots(st, cfindex, data=df, yvar="deaths")
  print(f$figp +  ggtitle(sprintf("Effect of no shutdown policies in %s",st)))

  figdev(sprintf("%s/tex/tables_and_figures/%s-index-dgrowth_deaths.pdf",rootdir,st), width=5,height=3.75)
  print(f$figdp)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-index-dcases_deaths.pdf",rootdir,st), width=5,height=3.75)
  print(f$figp)
  dev.off()
}

figs <- nationalplots(indexall, cfindex, "no shutdown policies", yvar="deaths")

figdev(paste(rootdir,"tex/tables_and_figures/us-index-dgrowth_deaths.pdf",sep="/"), width=5,height=3.75)
figs$g
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-index-dcases_deaths.pdf",sep="/"), width=5,height=3.75)
figs$c
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-index-rel_deaths.pdf",sep="/"), width=5,height=3.75)
figs$r
dev.off()

for (f in figs) print(f)
```
