---
title       : "Causal impact of masks, policies, behavior on early COVID-19 pandemic in the U.S."
subtitle    : "Counterfactuals"
author      :  Victor Chernozhukov, Paul Schrimpf, Hiro Kasahara
job         :
date        : "`r format(Sys.time(), '%d %B, %Y')`"
bibliography: "covid.bib"
link-citations: true
always_allow_html: true
output      :
    html_document :
        toc : true
        toc_depth : 3
        toc_float : true
        number_sections : true
        #theme : journal
        #css : 628notes.css
        code_folding: hide
        #lib_dir : deps
        self_cononontained : true
        fig_width: 8
        fig_height: 6
---

 <!-- To create html files from this, in R enter  -->
 <!-- library(rmarkdown) -->
 <!-- rmarkdown::render("regressionsWithPlots.Rmd") -->
 <!-- To create tex files from this, in R enter ` -->
 <!-- library(rmarkdown) -->
 <!-- rmarkdown::render("regressionsWithPlots.Rmd", output_format=latex_fragment()) -->


 <!-- to create the gh-pages hosted website in R enter -->
 <!-- rmarkdown::render_site() -->

 <!-- Then in a shell  -->
 <!-- $ ghp-import -p -n -m "$(date)" _site -->


```{r, include=FALSE}
options(knitr.table.format = function() {
  if (knitr::is_latex_output()) 'latex' else 'pandoc'
})
if (knitr::is_latex_output()) {
  knitr::opts_chunk$set(warning=FALSE, message=FALSE, echo=FALSE)
}
```

# Data Preparation
```{r setup, cache=FALSE, warning=FALSE, include=FALSE}

library(lfe)
library(stargazer)
library(knitr)
library(plm)
library(latex2exp)
library(ggplot2)
library(ggthemes)
library(estimatr)
library(gridExtra)
library(grid)
library(mvtnorm)
library(kableExtra)
library(data.table)
library(reshape2)

sgtype <- opts_knit$get("rmarkdown.pandoc.to")
sgstyle <- 'default'
colors <-  scale_color_solarized
colors_fill <- scale_fill_solarized
rootdir <- system("git rev-parse --show-toplevel", intern=TRUE)[1]
source(paste(rootdir,"cases_and_policies/R/dataprep.R",sep="/"))
source(paste(rootdir,"cases_and_policies/rmd/varlabels.R", sep="/"))
source(paste(rootdir,"cases_and_policies/rmd/regprep.R", sep="/"))
source(paste(rootdir,"cases_and_policies/rmd/generatetables.R", sep="/"))
source(paste(rootdir,"cases_and_policies/rmd/bootstrap_felm.R", sep="/"))

datelims <- c(as.Date("2020-03-07"), max(df$date))
```

# Counterfactuals


```{r cffuncs, cache=FALSE}
#' Given an estimated felm model with lags and differences, returns a
#' function that simulates the model.
#'
#' @param m estimated felm model
#' @param df Data
#' @param dvar Name of variable repressenting differences in the
#' model. dvar should be a difference of length difflag.
#' @param var Name of varaible representing levels in the model.
#' @param difflag difference length for dvar
#' @param lhsdiff if true the outcome is also dvar with the same
#' difflag, if not the outcome is var
#'
#' @return A function with arguments idvar==value of identifier for
#'   which to simulate and shockgen=function(n) that generates values
#'   for residuals. The default value of shockgen is to resample model
#'   residuals with replacement. The function returns a simulated path
#'   var beginning from the initial values for id==idvar and using the
#'   exogenous covariates of id==idvar.
#'
#' Example:
#' df <- df[order(df$id,$df$t),]
#' df$dy <- paneldiff(df$y, df$id, df$t, lag=2)
#' fmla <- dy ~ lag(y,3) + lag(dy,1) + x
#' # hack to get get id
#'
#' sim <- createforwardsim(fmla, df, "dy", "y", 2, TRUE)
#' # simulate with errors set to 0's
#' yhat <- sim(df$id[1], function(n) fill(0, n))
#' #' # simulate sampling residuals with replacement
#' yhat <- sim(df$id[1])
createforwardsim <- function(m, df, id="state",
                             dvar="dlogdc", var="logdc",
                             difflag=7, lhsdiff=TRUE) {
  m$coef[is.nan(m$coef)] <- 0

  # get index of observations used in mdl
  df$idxcfs <- 1:nrow(df)
  mid <- felm(update.felm.formula(m$formula, idxcfs ~ .), data=df, keepX=TRUE)
  idx <- mid$response

  # This  regexp will match dvar if preceded by start of string or (
  # and followed by end of string or ,
  di <- grepl(sprintf("(^|\\()%s(,|$)",dvar), rownames(m$coef))
  li <- grepl(sprintf("(^|\\()%s(,|$)",var), rownames(m$coef))
  dc <- m$coef[di]
  names(dc) <- rownames(m$coef)[di]
  lc <- m$coef[li]
  names(lc) <- rownames(m$coef)[li]
  # exogenous part of model
  if (!is.null(getfe(m))) {
    stop("createforwardsim not implemented with fixed effects. Include factors in the formula instead")
  }

  xt <- mid$X %*% (m$coef * as.numeric(!(di | li)))
  #xtold <- m$fitted.values - m$X %*% (m$coef * as.numeric(di | li))

  ## convert into AR model
  llags <- as.numeric(gsub("lag\\(.+, (\\d+)\\)","\\1", names(lc)))
  dlags <- as.numeric(gsub("lag\\(.+, (\\d+)\\)","\\1", names(dc)))
  maxlags <- max(c(llags, dlags+difflag, ifelse(lhsdiff, difflag, 0)))
  ar <- rep(0, maxlags)
  if (lhsdiff) ar[difflag]  <- ar[difflag] + 1
  ar[llags]  <- ar[llags] + lc
  ar[dlags] <- ar[dlags] +dc
  ar[dlags+difflag] <- ar[dlags+difflag]-dc

  if (!is(df, "pdata.frame")) {
    warning("df is not a pdata.frame. This function assumes df is sorted by id, time, and within each id, time is consecutive. Results will be incorrect if these assumptions are false.")
  } else {
    stopifnot(all(is.pconsecutive(df)))
  }

  sim <- function(idvar, shockgen=function(n=1) { sample(m$residuals, size=n,
                                                         replace=TRUE)},
                  samplecoefs=FALSE, nsim=100) {
    ids <- df[idx,id]
    if (samplecoefs) {
      V <- vcov(m)
      nac <- is.na(m$coefficients)
      e <- rep(0,nrow(V))
      Vnona <- V[!nac,!nac]
      e[!nac] <- t(rmvnorm(1, mean=rep(0, nrow(Vnona)), sigma=Vnona))

      dc <- m$coef[di] + e[di]
      lc <- m$coef[li] + e[li]
      ## convert into AR model
      ar <- rep(0, maxlags)
      if (lhsdiff) ar[difflag]  <- ar[difflag] + 1
      ar[llags]  <- ar[llags] + lc
      ar[dlags] <- ar[dlags] +dc
      ar[dlags+difflag] <- ar[dlags+difflag]-dc

      x <- xt[ids==idvar] + (mid$X[ids==idvar,] %*% (e * as.numeric(!(di | li))))
    } else {
      x <- xt[ids==idvar]
    }

    date <- df$date[idx[df[idx,id]==idvar]]
    names(date) <- NULL
    it0 <- which(df$date==date[1] & df[,id]==idvar)
    Y0 <- as.vector(sapply(1:(maxlags), function(k) lag(df[,var], k)[it0]))
    dY0 <- as.vector(sapply(1:(maxlags), function(k) lag(df[,dvar], k)[it0]))
    lagy0 <- Y0-dY0
    check <- lagy0[1:(length(Y0)-difflag)]- Y0[(difflag+1):length(Y0)]
    stopifnot(all(abs(check)<1e-8, na.rm=TRUE))
    na0 <- is.na(Y0[(difflag+1):length(Y0)])
    Y0[(difflag+1):length(Y0)][na0] <- lagy0[1:(length(Y0)-difflag)][na0]

    yt <- matrix(0,nrow=nsim,ncol=length(x))
    e <- matrix(shockgen(length(x)*nsim),nrow=nsim)
    Y0 <- t(matrix(rep(Y0,nsim),ncol=nsim))
    Y <- Y0
    for (t in 1:length(x)) {
      yt[,t] <- x[t] + Y %*% ar + e[,t]
      #Y <- c(yt[t], Y[-length(Y)])
      Y[,2:ncol(Y)] <- Y[,1:(ncol(Y)-1)]
      Y[,1] <- yt[,t]
    }
    return(list(y=cbind(Y0[, ncol(Y0):1],yt),
                date=c(date[1]-((maxlags):1),date)))
  }
}

meanandci <- function(yi, cfyi, date, p=0.9, difflag=7) {

  y <- colMeans(yi)
  cfy <- colMeans(cfyi)
  dy <- matrix(NA,nrow=nrow(y),ncol=ncol(y))
  dy[(difflag+1):nrow(y),]  <- diff(y,difflag)

  dcfy <- matrix(NA,nrow=nrow(cfy),ncol=ncol(cfy))
  dcfy[(difflag+1):nrow(cfy),]  <- diff(cfy,difflag)
  statframe <- function(y,suffix) {
    d <- data.frame(m=rowMeans(y, na.rm=TRUE),
                    cl=apply(y, 1, function(x) quantile(x, (1-p)/2,na.rm=TRUE)),
                    ch=apply(y, 1, function(x) quantile(x,1-(1-p)/2,na.rm=TRUE)))
    names(d) <- paste(names(d), suffix, sep="")
    d
  }

  cbind(data.frame(date=date),
        statframe(colMeans(exp(yi)),"y"), statframe(dy,"dy"), statframe(colMeans(exp(cfyi)),"cf"),
        statframe(dcfy,"dcf"), statframe(colMeans(exp(cfyi)-exp(yi)),"p"),
        statframe(dcfy-dy,"dp"))

}

nopstatesim <- function(st, nsim=200, simobs, simcf, data, var="logdc", dvar="dlogdc") {
  seed <- .Random.seed
  set.seed(seed)
  yi <- replicate(nsim, simobs(st, samplecoefs=TRUE, nsim=nsim/2)$y)
  set.seed(seed)
  cfyi <- replicate(nsim, simcf(st, samplecoefs=TRUE, nsim=nsim/2)$y)
  date <- simobs(st, samplecoefs=TRUE, nsim=2)$date
  est <- meanandci(yi,cfyi, date)
  est <- merge(est, data.frame(state=st,
                        obs=exp(data[data$state==st, var]),
                        dobs=data[data$state==st, dvar],
                        date=data[data$state==st,]$date),
               by="date", all=TRUE)
  return(list(edf=est, y=yi, cfy=cfyi))
}


cfplots <- function(st, cfdata=NULL, simobs=NULL, simcf=NULL,
                    yvar="cases", data) {
  xlims <- datelims
  pds <-
    c("Mandate.face.mask.use.by.employees.in.public.facing.businesses",
      "Date.closed.K.12.schools",
      "Stay.at.home..shelter.in.place",
      "Closed.movie.theaters",
      "Closed.restaurants.except.take.out",
      "Closed.non.essential.businesses")
  if (is.null(cfdata)) {
    est <- nopstatesim(st, simobs=simobs, simcf=simcf, data=data)
  } else {
    est <- subset(cfdata, cfdata$state==st)
  }

  dt  <- c()
  labels <- c()
  for (lbl in pds) {
    d <- unique(data[data$state==st,lbl])
    i <- which(dt==d)
    if (length(i)==0) {
      if (length(dt)==0)
        dt <- d
      else
        dt <- c(dt,d)
      labels <- c(labels,lbl)
    } else {
      labels[i]  <- paste(labels[i],lbl,sep=", ")
    }
  }
  labels <- labels[!is.na(dt)]
  dt <- dt[!is.na(dt)]

  figl <-
    ggplot(data=est, aes(x=date, y=obs)) +
    geom_line(aes(color="Observed", fill="Observed"), size=1.5) +
    geom_line(aes(y=my,color="Estimated"), size=1.2) +
    #geom_ribbon(aes(ymin=cly, ymax=chy, fill="Estimated"),
    #            alpha=0.3, show.legend=FALSE) +
    geom_line(aes(y=mcf, color="Counterfactual"), size=1.2) +
    #geom_ribbon(aes(ymin=clcf, ymax=chcf, fill="Without Policies"),
    #            alpha=0.3, show.legend=FALSE) +
    figtheme + colors() + colors_fill() +
    ylab(sprintf("%s in past week",yvar)) + labs(color="")  +
    theme(legend.position=c(0.2, 0.8)) +
    ggtitle(TeX(sprintf("%s in past week",yvar))) + xlim(xlims)

  figp <- ggplot(data=est, aes(x=date, y=mp)) +
    geom_line(size=1.2) +
    geom_ribbon(aes(ymin=clp, ymax=chp), alpha=0.3) +
    ylab(sprintf("Change in %s in past week", yvar)) + figtheme + colors() + colors_fill() +
    ggtitle(TeX(sprintf("Change in %s",yvar))) +
    xlim(xlims) +
    annotate("text",x=dt, y=rep(0.01,length(dt)),
             label=gsub("\\."," ", labels), size=4, angle=90, hjust=0)


  figd <- ggplot(data=est, aes(x=date, y=dobs)) +
    geom_line(aes(y=mdcf, color="Counterfactual"), size=1.5) +
    geom_line(aes(color="Observed", fill="Observed"), size=1.5) +
    geom_line(aes(y=mdy,color="Estimated"), size=1.5) +
    #geom_ribbon(aes(ymin=cldy, ymax=chdy, fill="Estimated"),
    #            alpha=0.3, show.legend=FALSE) +
    #geom_ribbon(aes(ymin=cldcf, ymax=chdcf, fill="Without Policies"),
    #            alpha=0.3, show.legend=FALSE) +
    figtheme + colors() + colors_fill() +
    ylab(TeX(sprintf("$\\Delta \\log \\Delta %s$",toupper(substring(yvar,1,1))))) +
    labs(color="")  +
    theme(legend.position=c(0.8, 0.8)) +
    ggtitle(TeX(sprintf("$\\Delta \\log \\Delta %s$",toupper(substring(yvar,1,1))))) +
    xlim(xlims)

  figdp <- ggplot(data=est, aes(x=date, y=mdp)) +
    geom_line(size=1.2) +
    geom_ribbon(aes(ymin=cldp, ymax=chdp), alpha=0.3) +
    ylab(TeX(sprintf("Change in $\\Delta \\log \\Delta %s$",toupper(substring(yvar,1,1))))) +
    figtheme + colors() + colors_fill() +
    ggtitle(TeX(sprintf("Change in $\\Delta \\log \\Delta %s$",toupper(substring(yvar,1,1))))) +
    xlim(xlims) +
    annotate("text",x=dt, y=rep(0.01,length(dt)),
             label=gsub("\\."," ", labels), size=4, angle=90, hjust=0)

  return(list(figd=figd, figdp=figdp, figl=figl, figp=figp))
}

# Extract E[log(dc) |state, t, parameters] with parameters ~ N(m$coef, vcov(m))
getmeans <- function(sl) {
  date <- sl$edf$date[!is.na(sl$edf$my)]
  state <- sl$edf$state[!is.na(sl$edf$state)]
  stopifnot(length(unique(state))==1)
  y0 <- apply(exp(sl$y), c(2,3), mean)
  long <- melt(y0)
  names(long)[names(long)=="value"] <- "meany0"

  log0 <- apply(sl$y, c(2,3), mean)
  long1 <- melt(log0)
  names(long1)[names(long1)=="value"] <- "meanlog0"
  long <- merge(long, long1, by=c("Var1","Var2"))

  y1 <- apply(exp(sl$cf), c(2,3), mean)
  long1 <- melt(y1)
  names(long1)[names(long1)=="value"] <- "meany1"
  long <- merge(long,long1, by=c("Var1","Var2"))

  log1 <- apply(sl$cf, c(2,3), mean)
  long1 <- melt(log1)
  names(long1)[names(long1)=="value"] <- "meanlog1"
  long <- merge(long, long1, by=c("Var1","Var2"))

  long$date <- date[long$Var1]
  long$state <- unique(state)
  long$sidx <- long$Var2
  long$Var1 <- NULL
  long$Var2 <- NULL
  return(long)
}

```

```{r statesim, cache=FALSE}
infovars <- list(c("dlogdc", "logdc")) #,
#c("dlogdc", "logdc"))
#infovars2 <- list(c( "logdc"))
tvars <- "dlogtests"
xlist <-list("")
interactions <-list("month", statevars)
ilist <- list(interactions) #, interactions)
iv <- list("0") #,"(testratedc ~ lag(testratedc,14))")

L <- 14
sdf <- subset(df, df$date>=datelims[1]) #+L)
regs <- mainregressions(sdf,
                        "dlogdc", pols, bvars, infovars, tvars, xlist, ilist, iv, L=L)
m <- regs$piy[[1]]$reg
cnames <- c(pols, infovars[[1]])
di <- dieff_table(lapply(regs$pib[[1]][1:4], function(x) x$reg$coef[,1]),
                  regs$pbiy[[1]]$reg$coef[,1],
                  regs$piy[[1]]$reg$coef[,1],
                  policies=cnames)

pib <- lapply(regs$pib[[1]][1:4], function(x) x$reg)
pbiy <- regs$pbiy[[1]]$reg
piy <- regs$piy[[1]]$reg

S <- 999
bs <- bootstrap_felm(sdf, c(pib,list(pbiy,piy)), S=S)
bdi <- lapply(1:S, function(i)
  dieff_table(lapply(1:4, function(j) bs[[j]][,i]),
              bs[[5]][,i], bs[[6]][,i],
              policies=cnames)
  )

subcoefs <- function(di, coef) {
  for (r in 1:(nrow(di)-1)) {
    i <- grep(rownames(di)[r] , rownames(coef))
    coef[i,1] <- di[r,"Average"]
  }
  return(coef)
}

bcoefs <- t(do.call(cbind, lapply(1:S, function(i) {
  coef <- matrix(bs[[6]][,i],ncol=1)
  rownames(coef) <- rownames(bs[[6]])
  return(subcoefs(bdi[[i]],coef ))
})))

V <- cov(bcoefs)
m$coefficients <- subcoefs(di, m$coefficients)
m$vcv <- V
m$clustervcv <- V
m$robustvcv <- V

simmodel <- createforwardsim(m, sdf)

# remove policies
cf <- sdf
cf[,pols] <- 0
nopmodel <- createforwardsim(m, cf)

rerunsim=FALSE
if (!file.exists("cfnop.Rda") || rerunsim) {
  cfnoplist <- lapply(unique(sdf$state), function(s)
    nopstatesim(s,simobs=simmodel,simcf=nopmodel,data=sdf))
  allmeans <- rbindlist(lapply(cfnoplist, getmeans))
  save(cfnoplist,allmeans, file="cfnop.Rda")
}
load("cfnop.Rda")
cfnop <- do.call(rbind, lapply(cfnoplist, function(x) x$edf))
```

### Removing Policies

We now examine the impact changing from the observed policies to
none. For illustrative purpose, we being by focusing on
Washington. Figure \ref{fig:WA-nop-dlogdc} shows the observed,
estimated average, and counterfactual without policies average of
$\Delta \log \Delta C$ in Washington. To compute the estimated and
counterfactual paths we use the estimates from column 2 of table
\ref{tab:policy-cases}. We set initial $\Delta \log \Delta C$ and
$\log \Delta C$ to their values first observed in the state we are
simulating. We hold all other regressors at their observed
values. Error terms are drawn with replacement from the residuals. We
do this many times and report the average over draws of the
residuals. To calculate confidence intervals, we repeat the above with
coefficients drawn randomly from their asymptotic distribution. The
shaded region is a pointwise 90\% confidence interval.


```{r WA-nop-dlogdc, cache=TRUE,fig.cap="Case growth without policies in Washington \\label{fig:WA-nop-dlogdc}"}
st <- "Washington"
f <- cfplots(st, cfnop, data=df)
grid.arrange(f$figd, f$figdp, nrow=2, top =
                                        textGrob(st,gp=gpar(fontsize=20,font=2)))

```

```{r WA-nop-cases, cache=TRUE, fig.cap="Cases without policies in Washington \\label{fig:WA-nop-cases}"}
grid.arrange(f$figl, f$figp, nrow=2) #,
#top = textGrob(st,gp=gpar(fontsize=20,font=2)))
```

# ```{r otherstates, cache=TRUE, fig.cap="Effect of Removing Policies"}
# if (!knitr::is_latex_output()) {
#   for( st in c("New York", "Florida", "Michigan", "Georgia", "Massachusetts")) {
#     f <- cfplots(st, cfnop, data=df)
#     grid.arrange(f$figd, f$figdp, nrow=2, top =  textGrob(st,gp=gpar(fontsize=20, font=2)))
#     grid.arrange(f$figl, f$figp, nrow=2, top =  textGrob(st,gp=gpar(fontsize=20, font=2)))
#   }
# }
# ```

# ```{r nop-state-figures, cache=TRUE, warning=FALSE}
# states <- c("Washington","Illinois","Massachusetts")
# for(st in states) {
#   f <- cfplots(st, cfnop, data=df)
# 
#   figdev(sprintf("%s/tex/tables_and_figures/%s-nop-dgrowth_v1.pdf",rootdir,st), width=5,height=3.75)
#   print(f$figdp)
#   dev.off()
# 
#   figdev(sprintf("%s/tex/tables_and_figures/%s-nop-growth_v1.pdf",rootdir,st), width=5,height=3.75)
#   print(f$figd + ylim(c(-1,5)))
#   dev.off()
# 
#   figdev(sprintf("%s/tex/tables_and_figures/%s-nop-cases_v1.pdf",rootdir,st), width=5,height=3.75)
#   print(f$figl)
#   dev.off()
# 
#   figdev(sprintf("%s/tex/tables_and_figures/%s-nop-dcases_v1.pdf",rootdir,st), width=5,height=3.75)
#   print(f$figp)
#   dev.off()
# }
# ```

## National

These figures show the average effect on $\Delta \log \Delta C$ across
states, and the total national effect on cases in past week.

```{r nop-all, cache=TRUE}
abbrv <- unique(df[,c("state","ST")])
cfnop <- merge(cfnop, abbrv, by="state")

nationalplots <- function(alldf,cfdf, cfname, yvar="cases") {
  alldf[order(state, sidx, date), dlog1:=c(rep(NA,7),diff(meanlog1,7)), by=.(state, sidx)]
  alldf[order(state, sidx, date),dlog0:=c(rep(NA,7),diff(meanlog0,7)), by=.(state, sidx)]
  national <- alldf[ , .(diff=sum(meany1-meany0), y0=sum(meany0), dlog=mean(dlog1-dlog0)), by=.(date, sidx)]

  adf <- national[, .(mp=mean(diff),
                      clo=quantile(diff, 0.05),
                      chi=quantile(diff, 0.95),
                      rel=mean(diff/y0),
                      rlo=quantile(diff/y0, 0.05),
                      rhi=quantile(diff/y0, 0.95),
                      mlog=mean(dlog, na.rm=TRUE),
                      llo=quantile(dlog, 0.025, na.rm=TRUE),
                      lhi=quantile(dlog, 0.975, na.rm=TRUE)
                      ),
                  keyby=.(date)]

  figdlog <- ggplot(cfdf, aes(x=date)) +
    #geom_text(aes(y=mdp, label=ST),alpha=0.2) +
    geom_point(aes(y=mdp),alpha=0.2, size=0.5) +
    xlim(datelims) +
    geom_line(data=adf,  aes(x=date, y=mlog, color=""), size=2) + figtheme +
    geom_ribbon(data=adf, aes(ymin=llo, ymax=lhi, fill=""),alpha=0.3) +
    colors() + theme(legend.position="none") +
    ylab(TeX(sprintf("Change in $\\Delta \\log \\Delta %s$",
                     toupper(substring(yvar,1,1))
                     ))) +
    colors_fill() +
     ggtitle(sprintf("Effect of %s on %s growth",cfname, substring(yvar,1,nchar(yvar)-1)))

  figc <- ggplot(cfdf, aes(x=date)) +
    #geom_text(aes(y=mp, label=ST),alpha=0.5) +
    geom_point(aes(y=mp),alpha=0.2, size=0.5) +
    xlim(datelims) +
    geom_line(data=adf, aes(x=date, y=mp, color=""), size=2) +
    figtheme +
    geom_ribbon(data=adf, aes(ymin=clo, ymax=chi, fill=""),alpha=0.3) +
    colors() + theme(legend.position="none") +
    ylab(sprintf("%s in the past week", yvar)) +
    colors_fill() +
    ggtitle(sprintf("Effect of %s on %s",cfname, yvar))

  figr <- ggplot(adf, aes(x=date, y=rel)) +
    xlim(datelims) +
    geom_line(aes(color=""), size=2) + figtheme +
    geom_ribbon(data=adf, aes(ymin=rlo, ymax=rhi, fill=""),alpha=0.3) +
    colors() + theme(legend.position="none") +
    ylab(sprintf("relative increase in %s",yvar)) +
    colors_fill() +
    ggtitle(sprintf("Relative effect of %s",cfname))
  return(list(g=figdlog, c=figc, r=figr))
}

figs <- nationalplots(allmeans,cfnop, "removing policies")

figdev(paste(rootdir,"tex/tables_and_figures/us-nop-dgrowth_v1.pdf",sep="/"), width=5,height=3.75)
figs$g
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-nop-dcases_v1.pdf",sep="/"), width=5,height=3.75)
figs$c
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-nop-rel_v1.pdf",sep="/"), width=5,height=3.75)
figs$r
dev.off()

for (f in figs) print(f)
```

## Non-Essential Business Closures

```{r business_sim, cache=FALSE}
# remove non essential business closure
cf <- sdf
cf[,"pnonessential"] <- 0
nobmodel <- createforwardsim(m, cf)

if (!file.exists("cfnb.Rda") || rerunsim) {
  cfnblist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=nobmodel, data=sdf))
  nball <- rbindlist(lapply(cfnblist, getmeans))
  save(cfnblist,nball, file="cfnb.Rda")
}
load("cfnb.Rda")
cfnb <- do.call(rbind, lapply(cfnblist, function(x) x$edf))
```

```{r, cache=TRUE}
cfnb <- merge(cfnb, abbrv, by="state")
for (st in c("Washington","Illinois","Massachusetts")) {
  f <- cfplots(st, cfnb, data=df)
  f$figp + ggtitle(sprintf("Effect of never shutting down non-essential businesses in %s",st))
  #grid.arrange(f$figd + ggtitle(), f$figdp + ggtitle(), nrow=2, top =
  #textGrob(st,gp=gpar(fontsize=20,font=2)))

  figdev(sprintf("%s/tex/tables_and_figures/%s-nb-dgrowth_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figdp)
  dev.off()


  figdev(sprintf("%s/tex/tables_and_figures/%s-nb-dcases_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figp)
  dev.off()
}


figs <- nationalplots(nball, cfnb,
                      "never shutting down non-essential\nbusinesses")

figdev(paste(rootdir,"tex/tables_and_figures/us-nb-dgrowth_v1.pdf",sep="/"), width=5,height=3.75)
figs$g
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-nb-dcases_v1.pdf",sep="/"), width=5,height=3.75)
figs$c
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-nb-rel_v1.pdf",sep="/"), width=5,height=3.75)
figs$r
dev.off()

for (f in figs) print(f)
```

## Masks

We now look at what happens when all states impose a mask mandate on
April 1st.

```{r mask_sim, cache=FALSE}
cf <- df
cf[,"Mandate.face.mask.use.by.employees.in.public.facing.businesses"]  <- as.Date("2020-04-01")
cf[,"pmaskbus"] <-
  smoothedpolicy(cf,"Mandate.face.mask.use.by.employees.in.public.facing.businesses",type="ma",bw=6)
cf <- subset(cf, cf$date>=datelims[1]) #+L)
maskmodel <- createforwardsim(m, cf)

if (!file.exists("cfmask.Rda") || rerunsim) {
  cfmasklist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=maskmodel, data=sdf))
  maskall <- rbindlist(lapply(cfmasklist, getmeans))
  save(cfmasklist, maskall, file="cfmask.Rda")
}
load("cfmask.Rda")
cfmask <- do.call(rbind, lapply(cfmasklist, function(x) x$edf))
```

```{r, cache=TRUE}
cfmask <- merge(cfmask, abbrv, by="state")
for (st in c("Washington","Illinois","Massachusetts")) {
  f <- cfplots(st, cfmask, data=df)
  print(f$figp +  ggtitle(sprintf("Effect of mandating masks on April 1st in %s",st)))

  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-dgrowth_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figdp)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-dcases_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figp)
  dev.off()
}

figs <- nationalplots(maskall, cfmask, "mandating masks on April 1st\n")

figdev(paste(rootdir,"tex/tables_and_figures/us-mask-dgrowth_v1.pdf",sep="/"), width=5,height=3.75)
figs$g
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-mask-dcases_v1.pdf",sep="/"), width=5,height=3.75)
figs$c
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-mask-rel_v1.pdf",sep="/"), width=5,height=3.75)
figs$r
dev.off()

for (f in figs) print(f)
```

## Stay at home

We now look at what happens if stay at home orders were never issued.

```{r home, cache=FALSE}
cf <- df
cf[,"pshelter"] <- 0
cf <- subset(cf, cf$date>=datelims[1]) #+L)
noshelter <- createforwardsim(m, cf)

if (!file.exists("cfshelter.Rda") || rerunsim) {
  cfshelterlist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=noshelter, data=sdf))
  shelterall <- rbindlist(lapply(cfshelterlist, getmeans))
  save(cfshelterlist, shelterall, file="cfshelter.Rda")
}
load("cfshelter.Rda")
cfshelter <- do.call(rbind, lapply(cfshelterlist, function(x) x$edf))
```

```{r, cache=TRUE}
cfshelter <- merge(cfshelter, abbrv, by="state")
for (st in c("Washington","Illinois","Massachusetts")) {
  f <- cfplots(st, cfshelter, data=df)
  print(f$figp +  ggtitle(sprintf("Effect of no stay at home order in %s",st)))

  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dgrowth_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figdp)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dcases_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figp)
  dev.off()
}

figs <- nationalplots(shelterall, cfshelter, "no stay at home orders")

figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-dgrowth_v1.pdf",sep="/"), width=5,height=3.75)
figs$g
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-dcases_v1.pdf",sep="/"), width=5,height=3.75)
figs$c
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-rel_v1.pdf",sep="/"), width=5,height=3.75)
figs$r
dev.off()

for (f in figs) print(f)
```

# Deaths

```{r statesim-deaths, cache=FALSE}
infovars <- list(c("dlogdd", "logdd")) #,
#c("dlogdc", "logdc"))
#infovars2 <- list(c( "logdc"))
tvars <- NULL
xlist <-list("")
interactions <-list("month", statevars)
ilist <- list(interactions) #, interactions)
iv <- list("0") #,"(testratedc ~ lag(testratedc,14))")

L <- 21
sdf <- subset(df, df$date>=datelims[1]) #+L)
regs <- mainregressions(sdf,
                        "dlogdd", pols, bvars, infovars, tvars, xlist, ilist, iv, L=L)

m <- regs$piy[[1]]$reg

cnames <- c(pols, infovars[[1]])
di <- dieff_table(lapply(regs$pib[[1]][1:4], function(x) x$reg$coef[,1]),
                  regs$pbiy[[1]]$reg$coef[,1],
                  regs$piy[[1]]$reg$coef[,1],
                  policies=cnames)

pib <- lapply(regs$pib[[1]][1:4], function(x) x$reg)
pbiy <- regs$pbiy[[1]]$reg
piy <- regs$piy[[1]]$reg

S <- 999
bs <- bootstrap_felm(sdf, c(pib,list(pbiy,piy)), S=S)
bdi <- lapply(1:S, function(i)
  dieff_table(lapply(1:4, function(j) bs[[j]][,i]),
              bs[[5]][,i], bs[[6]][,i],
              policies=cnames)
  )

bcoefs <- t(do.call(cbind, lapply(1:S, function(i) {
  coef <- matrix(bs[[6]][,i],ncol=1)
  rownames(coef) <- rownames(bs[[6]])
  return(subcoefs(bdi[[i]],coef ))
})))
V <- cov(bcoefs)
m$coefficients <- subcoefs(di, m$coefficients)
m$vcv <- V
m$clustervcv <- V
m$robustvcv <- V

simmodel <- createforwardsim(m, sdf, dvar="dlogdd", var="logdd")

# remove policies
cf <- sdf
cf[,pols] <- 0
nopmodel <- createforwardsim(m, cf, dvar="dlogdd", var="logdd")

rerunsim=FALSE
if (!file.exists("cfnop_deaths.Rda") || rerunsim) {
  cfnoplist <- lapply(unique(df$state), function(s)
    nopstatesim(s,simobs=simmodel,simcf=nopmodel, data=sdf, var="logdd", dvar="dlogdd"))
  allmeans <- rbindlist(lapply(cfnoplist, getmeans))
  save(cfnoplist,allmeans, file="cfnop_deaths.Rda")
}
load("cfnop_deaths.Rda")
cfnop <- do.call(rbind, lapply(cfnoplist, function(x) x$edf))
```

### Removing Policies


```{r WA-nop-dlogdd, cache=TRUE,fig.cap="Case growth without policies in Washington \\label{fig:WA-nop-dlogdc}"}
st <- "Washington"
f <- cfplots(st, cfnop, yvar="deaths", data=df)
grid.arrange(f$figd, f$figdp, nrow=2, top =
                                        textGrob(st,gp=gpar(fontsize=20,font=2)))

```

```{r WA-nop-deaths, cache=TRUE, fig.cap="Cases without policies in Washington \\label{fig:WA-nop-cases}"}
grid.arrange(f$figl, f$figp, nrow=2) #,
#top = textGrob(st,gp=gpar(fontsize=20,font=2)))
```

# ```{r otherstates-deaths, cache=TRUE, fig.cap="Effect of Removing Policies"}
# if (!knitr::is_latex_output()) {
#   for( st in c("New York", "Florida", "Michigan", "Georgia", "Massachusetts")) {
#     f <- cfplots(st, cfnop, yvar="deaths", data=df)
#     grid.arrange(f$figd, f$figdp, nrow=2, top =  textGrob(st,gp=gpar(fontsize=20, font=2)))
#     grid.arrange(f$figl, f$figp, nrow=2, top =  textGrob(st,gp=gpar(fontsize=20, font=2)))
#   }
# }
```

# ```{r nop-state-figures-deaths, cache=TRUE, warning=FALSE}
# states <- c("Washington","Illinois","Massachusetts")
# for(st in states) {
#   f <- cfplots(st, cfnop, yvar="deaths", data=df)
# 
#   figdev(sprintf("%s/tex/tables_and_figures/%s-nop-dgrowth_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
#   print(f$figdp)
#   dev.off()
# 
#   figdev(sprintf("%s/tex/tables_and_figures/%s-nop-growth_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
#   print(f$figd + ylim(c(-1,5)))
#   dev.off()
# 
#   figdev(sprintf("%s/tex/tables_and_figures/%s-nop-cases_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
#   print(f$figl)
#   dev.off()
# 
#   figdev(sprintf("%s/tex/tables_and_figures/%s-nop-dcases_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
#   print(f$figp)
#   dev.off()
# }
# ```

## National


```{r nop-all-deaths, cache=TRUE}
abbrv <- unique(df[,c("state","ST")])
cfnop <- merge(cfnop, abbrv, by="state")

figs <- nationalplots(allmeans,cfnop, "removing policies", yvar="deaths")

figdev(paste(rootdir,"tex/tables_and_figures/us-nop-dgrowth_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$g
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-nop-dcases_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$c
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-nop-rel_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$r
dev.off()

for (f in figs) print(f)
```

## Non Essential Business Closures

```{r business_sim-deaths, cache=FALSE}
# remove non essential business closure
cf <- sdf
cf[,"pnonessential"] <- 0
nobmodel <- createforwardsim(m, cf, dvar="dlogdd", var="logdd")

if (!file.exists("cfnb_deaths.Rda") || rerunsim) {
  cfnblist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=nobmodel, data=sdf, var="logdd", dvar="dlogdd"))
  nball <- rbindlist(lapply(cfnblist, getmeans))
  save(cfnblist,nball, file="cfnb_deaths.Rda")
}
load("cfnb_deaths.Rda")
cfnb <- do.call(rbind, lapply(cfnblist, function(x) x$edf))
```

```{r, cache=TRUE}
cfnb <- merge(cfnb, abbrv, by="state")
for (st in c("Washington","Illinois","Massachusetts")) {
  f <- cfplots(st, cfnb, yvar="deaths", data=df)
  f$figp + ggtitle(sprintf("Effect of never shutting down non-essential businesses in %s",st))
  #grid.arrange(f$figd + ggtitle(), f$figdp + ggtitle(), nrow=2, top =
  #textGrob(st,gp=gpar(fontsize=20,font=2)))

  figdev(sprintf("%s/tex/tables_and_figures/%s-nb-dgrowth_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figdp)
  dev.off()


  figdev(sprintf("%s/tex/tables_and_figures/%s-nb-dcases_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figp)
  dev.off()
}


figs <- nationalplots(nball, cfnb,
                      "never shutting down non-essential\nbusinesses", yvar="deaths")

figdev(paste(rootdir,"tex/tables_and_figures/us-nb-dgrowth_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$g
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-nb-dcases_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$c
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-nb-rel_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$r
dev.off()

for (f in figs) print(f)
```

## Masks

We now look at what happens when all states impose a mask mandate on
April 1st.

```{r mask_sim-deaths, cache=FALSE}
cf <- df
cf[,"Mandate.face.mask.use.by.employees.in.public.facing.businesses"]  <- as.Date("2020-04-01")
cf[,"pmaskbus"] <-
  smoothedpolicy(cf,"Mandate.face.mask.use.by.employees.in.public.facing.businesses",type="ma",bw=6)
cf <- subset(cf, cf$date>=datelims[1]) #+L)
maskmodel <- createforwardsim(m, cf, dvar="dlogdd", var="logdd")

if (!file.exists("cfmask_deaths.Rda") || rerunsim) {
  cfmasklist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=maskmodel, data=sdf, var="logdd", dvar="dlogdd"))
  maskall <- rbindlist(lapply(cfmasklist, getmeans))
  save(cfmasklist, maskall, file="cfmask_deaths.Rda")
}
load("cfmask_deaths.Rda")
cfmask <- do.call(rbind, lapply(cfmasklist, function(x) x$edf))
```

```{r, cache=TRUE}
cfmask <- merge(cfmask, abbrv, by="state")
for (st in c("Washington","Illinois","Massachusetts")) {
  f <- cfplots(st, cfmask, yvar="deaths", data=df)
  print(f$figp +  ggtitle(sprintf("Effect of mandating masks on April 1st in %s",st)))

  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-dgrowth_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figdp)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-dcases_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figp)
  dev.off()
}

figs <- nationalplots(maskall, cfmask, "mandating masks on April 1st\n", yvar="deaths")

figdev(paste(rootdir,"tex/tables_and_figures/us-mask-dgrowth_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$g
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-mask-dcases_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$c
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-mask-rel_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$r
dev.off()

for (f in figs) print(f)
```

## Stay at home

We now look at what happens if stay at home orders were never issued.

```{r home-deaths, cache=FALSE}
cf <- df
cf[,"pshelter"] <- 0
cf <- subset(cf, cf$date>=datelims[1]) #+L)
noshelter <- createforwardsim(m, cf, dvar="dlogdd", var="logdd")

if (!file.exists("cfshelter_deaths.Rda") || rerunsim) {
  cfshelterlist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=noshelter, data=sdf))
  shelterall <- rbindlist(lapply(cfshelterlist, getmeans))
  save(cfshelterlist, shelterall, file="cfshelter_deaths.Rda")
}
load("cfshelter_deaths.Rda")
cfshelter <- do.call(rbind, lapply(cfshelterlist, function(x) x$edf))
```

```{r, cache=TRUE}
cfshelter <- merge(cfshelter, abbrv, by="state")
for (st in c("Washington","Illinois","Massachusetts")) {
  f <- cfplots(st, cfshelter, data=df, yvar="deaths")
  print(f$figp +  ggtitle(sprintf("Effect of no stay at home order in %s",st)))

  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dgrowth_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figdp)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dcases_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figp)
  dev.off()
}

figs <- nationalplots(shelterall, cfshelter, "no stay at home orders", yvar="deaths")

figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-dgrowth_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$g
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-dcases_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$c
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-rel_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$r
dev.off()

for (f in figs) print(f)
```
